Symbolic Weighted Language Models and Parsing over Infinite Alphabets

### 2021-05-06 main.tex
construction image transducer

@@

\begin{example}
** classical examples **
semirings      
\end{example}

\subsection{Label Theory}
\label{section:symbols}
 We shall define the functions labeling the transitions of Sw automata and transducers.
Let $\Sigma$ and $\Delta$ be two countable sets of symbols
called respectively input and output \emph{alphabets}, 
and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.

@@

\noindent
The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
are synthesized into a function~$\wei$ 
from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.
%
When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

\subsection{Properties}

\begin{proposition}
Given a \SWT $T$ 
over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
and $s \in \Sigma^*$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ such that for all $t \in \Delta^*$, 
$A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \vec{\wei}, \final >$ and let $n = |s|$, the length of $s$.

\qed
\end{proof}

The construction time and size for $A_{s, T}$ are $O(| s | . \| T \|)$.

@@

\section{Symbolic Weighted Parsing}

>>>>

% array and tabular
\usepackage{array}
\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

@@

\begin{example}
Figure~\ref{gif:semirings} presented examples of semirings interesting in practice 
and enjoying the above properties.
\end{example}

\begin{figure}[t]
\begin{center}
\begin{tabular}{|c|c|C{2em}|C{2em}|C{2em}|C{2em}|}
\hline
        & domain & $\oplus$ & $\otimes$ & $\zero$  & $\one$\\ %& nat. ordering\\
\hline\hline
Boolean & $\{\bot, \top\}$ & $\vee$ & $\wedge$ & $\bot$ & $\top$\\ %& $\top \leq_\oplus \bot$  \\
\hline
Counting & $\mathbb{N}$ & $+$ & $\times$ & 0 & 1 \\
\hline
Viterbi & $[0, 1] \subset \mathbb{R}$ & $\mathit{max}$ & $\times$ & 0 & 1\\ % & $x \leq_\oplus y \iff x \ge y$  \\
\hline
Tropical & $\mathbb{R}_+ \cup \{ +\infty\}$ & $\mathit{min}$ & $+$ & $+\infty$ & $0$\\ % & $x \leq_\oplus y \iff x \leq y$   \\
\hline
%MaxPlus & $\mathbb{Q} \cup \{ -\infty\}$ & $\mathsf{max}$ & $+$ & $-\infty$ & $0$ \\
%\hline
%Word lang. & $2^{\Sigma^*}$ & $\cup$ & $\cdot$ & $\emptyset$ & $\{ \varepsilon \}$ \\
%\hline
\end{tabular}
\end{center}
\vskip-1em
\caption{Some commutative, bounded, total and complete semirings.}
\label{fig:semirings}
\end{figure}

\subsection{Label Theory}
\label{section:symbols}
 We shall now define the functions labeling the transitions of SW-automata and -transducers.
Let $\Sigma$ and $\Delta$ be two countable sets of symbols
called respectively input and output \emph{alphabets}, 
and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.

@@

\noindent
The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
are synthesized into a function~$\wei$ 
from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$ 
as above.
%
When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

\subsection{Properties}
The two following properties will be useful to our approach on 
symbolic weighted parsing in Section~\ref{sec:parsing}.

\begin{proposition}
Given a \SWT $T$ 
over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
and $s \in \Sigma^*$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ such that for all $t \in \Delta^*$, 
$A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \vec{\wei}, \final >$ 
where $\vec{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
has an associated synthesized form $\wei$, 
and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
%$n = |s|$ be the length of $s$.
%
Let $Q' = [1..n] \times Q$ be the state set of $A_{T, s}$,
and let, for all $q \in Q$, 
$\init'(1, q) = \init(q)$ and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
and $\final'(n, q) = \final(q)$ and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.

\noindent
We define the functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$
in their synthesized form: for all $1 \leq i, j \leq n$, $q, q' \in Q$ and $a \in \Sigma$, 
and $b \in \Delta$, 
\[
\begin{array}{lclcl}
\wei'(\< i, q>, \epsilon, \< i, q'>)   & = & \wei(q, \epsilon, \epsilon, q'), & \quad & \\ 
\wei'(\< i, q>, \epsilon, \< i+1, q'>) & = & \wei(q, s_i, \epsilon, q') & & \mathrm{if~} i < n,\\ 
\wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1,\\ 
\wei'(\< i, q>, b, \< i, q'>)          & = & \wei(q, \epsilon, b, q'),\\ 
\wei'(\< i, q>, b, \< i+1, q'>)        & = & \wei(q, s_i, b, q') & & \mathrm{if~} i < n,\\ 
\wei'(\< i, q>, b, \< j, q'>)          & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1. 
\end{array}      
\]
Let $A_{T, s} = \< Q', \init', \vec{\wei'}, \final' >$.
\qed
\end{proof}

The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
where the size $\| T \|$ of $T$ is its number of states $|Q|$.

@@

\section{Symbolic Weighted Parsing}
\label{sec:parsing}
